---
title: Proposition 50 
format: dashboard
---

# Congressional Maps
## Row 

```{r}
#| include: false

# log all data for election_2025 data 
library(tidyverse)
library(sf)
SR_vote_data <- read_csv("../Downloads/data/SR_Precinct_Votes.csv")
SR_shapefiles <- st_read("../Downloads/data/shapefiles/SR_precinct_shapefiles")
redistrict_data <- st_read("../Downloads/data/shapefiles/redistrict")

#ensure same projection
SR_shapefiles <- SR_shapefiles |>
    st_transform(3310)|> # switch to equal area projection first
    st_set_precision(1) |> # snap points to 1 m grid
    st_make_valid() |> # fix self-intersections/bow-ties
    st_collection_extract("POLYGON")

#cleaning and combining SR_vote_data & SR_shapefiles 
join_data <- SR_vote_data|>
    select("SRPREC_KEY", "CDDIST", "CNGDEM01", "CNGDEM02", "CNGREP01", "CNGREP02")|>
    mutate(CNGDEM01 = str_replace_all(CNGDEM01, "[\\*]+", "0"), CNGDEM02 = str_replace_all(CNGDEM02, "[\\*]+", "0"), 
    CNGREP01 = str_replace_all(CNGREP01, "[\\*]+", "0"), CNGREP02 = str_replace_all(CNGREP02, "[\\*]+", "0"))|> #convert * to 0
    drop_na()|> # remove NAs
    mutate(CNGDEM01= as.numeric(CNGDEM01), CNGDEM02= as.numeric(CNGDEM02), CNGREP01= as.numeric(CNGREP01), CNGREP02= as.numeric(CNGREP02))|> #convert columns into numeric
    left_join(SR_shapefiles, by= join_by(SRPREC_KEY == SRPREC_KEY))|>
    mutate(DEMVOTE = sum(CNGDEM01, CNGDEM02), REPVOTE = sum(CNGREP01, CNGREP02))

#cleaning redistrict data 
redistrict_data <- redistrict_data|>
    select(DISTRICT, geometry)

#convert to join_data to sf 
join_data_sf <- st_sf(geometry = join_data$geometry, DEMVOTE= join_data$DEMVOTE, REPVOTE = join_data$REPVOTE)

#plot geometries 
district_2024 <- plot(st_geometry(join_data_sf))
district_2025 <- plot(st_geometry(redistrict_data))

#ensuring same crs/ projection
redistrict_data$geometry <- redistrict_data$geometry |>
    st_transform(3310) |> # switch to equal area projection first
    st_set_precision(1) |> # snap points to 1 m grid
    st_make_valid() |> # fix self-intersections/bow-ties
    st_collection_extract("POLYGON")

#area interpolate 
interp <- st_interpolate_aw(join_data_sf, redistrict_data, extensive = TRUE)
interp

```

```{r}
#| include: false

# run all data again for election_2024 data 
library(readxl)
precinct_votes <- read_csv("data/g24_sov_by_g24_svprec.csv")
# COUNTY, FIPS, CDIST, TOTVOTE, DEMVOTE, REPVOTE, CNGDEM01, CNGDEM02, CNGREP01,, CNGREP02
district_votes <- read_excel("data/g24-results-by-district.xlsx")
candidates <- read_csv("data/g24-candidates-by-district.csv")

precinct_votes <- precinct_votes|>
    select("COUNTY", "FIPS", "CDDIST", "TOTVOTE", "DEMVOTE", "REPVOTE",, "CNGDEM01", "CNGDEM02", "CNGREP01", "CNGREP02")|>
    drop_na()|> # remove NAs
    mutate(CNGDEM01= as.numeric(CNGDEM01),CNGDEM02= as.numeric(CNGDEM02), CNGREP01= as.numeric(CNGREP01), CNGREP02= as.numeric(CNGREP02)) 

election_2024 <- precinct_votes|>
    group_by(CDDIST)|>
    drop_na()|>
    summarize(total_dem = sum(CNGDEM01 + CNGDEM02), total_rep = sum(CNGREP01 + CNGREP02))|> #sum the total democratic and republican votes by district
    slice(-1)
```


```{r}
#| include: false

#using the join_data, calculate which party wins 

district_2024 <- st_read("../Downloads/2024_sh")

mapping_2024 <- district_2024|>
    mutate(CDDIST = 1:52)|>
    full_join(election_2024)|>
    select("CDDIST", "geometry", "total_dem", "total_rep")|>
    mutate(Party = ifelse(total_dem > total_rep, "Democrat", "Republican"))

#convert to sf 
mapping_2024_sf <- st_sf(CDDIST = mapping_2024$CDDIST, geometry = mapping_2024$geometry, party = mapping_2024$Party)
```

```{r}
#| echo: false
ggplot()+
    geom_sf(data = mapping_2024_sf, aes(fill= party))+ 
    scale_fill_manual( values = c("Democrat" = "light blue", "Republican" = "red"))+
    labs(title = "Congressional Results in 2024", fill = "Party Win")
```


## Row

```{r}
#| include: false
mapping_2025 <- interp|>
    mutate(Party = ifelse(DEMVOTE > REPVOTE, "Democrat", "Republican"))
```

```{r}
#| echo: false
ggplot()+
    geom_sf(data = mapping_2025, aes(fill= Party))+ 
    scale_fill_manual( values = c("Democrat" = "light blue", "Republican" = "red"))+
    labs(title = "Proposed Congressional Results in 2025 w/ the Passing of Prop 50", fill = "Party Win")
```

# Statistics
## Row 

Seats Won By Democrats, Mean Median Score, Efficiency Gap in 2024
```{r}
#| echo: false

#calculating mean_median score 
election_2024|>
    mutate(total = total_dem + total_rep)|>
    mutate(prop = (total_dem/ total))|>
    summarize(mean_median = mean(prop) - median(prop))

# proportion won by democrats 
election_2024|>
    summarize(Democratic_seats_Won = sum(total_dem > total_rep))
    
#calculating efficiency gap
n_wasted <- function(by, against){
    w_threshold <- floor((by +against)/2) +1 
    n_wasted <- ifelse(by > against, by - w_threshold, by)
    n_wasted 
}

election_2024 |>
    mutate(
    d_wasted = n_wasted(total_dem, total_rep), 
    r_wasted = n_wasted(total_rep, total_dem)
    ) |>
    summarize(
        total_d_wasted = sum(d_wasted), 
        total_r_wasted = sum(r_wasted), 
        total_votes = sum(total_dem, total_rep)
    )|>
    mutate(EG = (total_d_wasted - total_r_wasted)/total_votes )
```

## Row
Seats Won By Democrats, Mean Median Score, Efficiency Gap w/ Proposed Proposition 50 Redistricting
```{r}
#| echo: false
#calculating mean_median score 
interp|>
    mutate(total = DEMVOTE + REPVOTE)|>
    mutate(prop = (DEMVOTE/ total))|>
    summarize(mean_median = mean(prop) - median(prop))

#seats won by democrats 
interp|>
    summarize(Democratic_seats_Won = sum(DEMVOTE > REPVOTE))

#EG
interp |>
    mutate(
    d_wasted = n_wasted(DEMVOTE, REPVOTE), 
    r_wasted = n_wasted(REPVOTE, DEMVOTE)
    ) |>
    summarize(
        total_d_wasted = sum(d_wasted), 
        total_r_wasted = sum(r_wasted), 
        total_votes = sum(DEMVOTE, REPVOTE)
    )|>
    mutate(EG = (total_d_wasted - total_r_wasted)/total_votes )
```

# Methodology 
::: {.card}
I used data from the California Statewide website. I calculated the proposed redistricting wins by using area interpolation. Note that each of the statistics individually such as mean_median score and efficiency gap cannot tell the full story of gerrymandering. 
:::