---
title: Exploratory Data Analysis
format: typst
---

### Question 1

 How many precincts featured two candidates from the same party? 

### Answer 1

```{r}
question_1 <- district_votes|>
    filter((Democrat == "DEM" & Republican == "DEM"| Democrat == "REP" & Republican == "REP")) #districts who have both candidates from the same party running 

nrow(question_1) 

```

Four districts had candidates running from the same party. 

### Question 2

Which county had the greatest win? By how much?

### Answer 2

```{r}
question_2 <- precinct_votes|>
    group_by(COUNTY)|>
    summarize(difference = sum(CNGDEM01, CNGDEM02) - sum(CNGREP01, CNGREP02))|>
    mutate(difference = abs(difference))|>
    arrange(desc(difference))|>
    slice(1,)

question_2
```

County 19 had the greatest difference in votes which was by 9143708. 

### Question 3

Does the precinct votes when grouped together by district match the district votes? 

### Answer 3


```{r}
new_district_votes <- district_votes|>
    filter(!(Democrat == "DEM" & Republican == "DEM"))|> #don't want districts who have both democrat candidates running 
    filter(!(Democrat == "REP" & Republican == "REP"))|> #don't want districts who have both republican candidates running 
    filter(District == "District Totals")|> #only keep the district votes from each district |>
    mutate(District = c(1:52)) #numbered rows to represent each district

question_3 <- precinct_votes|>
    group_by(CDDIST)|>
    summarize(democrat_total = sum(CNGDEM01,CNGDEM02), republican_total = sum(CNGREP01,CNGREP02))|>
    slice(-1)|> #remove District 0 which represents the total 
    full_join(new_district_votes, by = join_by(CDDIST == District))
```

No they are off by a few digits.