---
title: "Practice Worksheet: Gerrymandering"
format: typst
---

Recall the 3 quantitative approaches to measuring gerrymandering discussed in lecture:\
1. Mean - Median Score (Grofman, 1983)\
2. Efficiency Gap (Stephanopolous and McGee, 2015)\
3. Ensembles / Extreme Outliers\
Use the following table to practice applying the first two techniques.\

| District | D Votes | R Votes | Result  |
|:---------:|:-------:|:-------:|:--------|
| 1 | 62 | 38 | D wins |
| 2 | 55 | 45 | D wins |
| 3 | 44 | 56 | R wins |
| 4 | 47 | 53 | R wins |
| 5 | 46 | 54 | R wins |
| **Total** | **254** | **246** |          |

```{r}
#| message: false
#| echo: false
library(tidyverse)
votes <- tibble(District = 1:5,
                d_votes = c(62, 55, 44, 47, 46),
                r_votes = c(38, 45, 56, 53, 54),
                result = c(rep("D wins", 2), rep("R wins", 3)))
```

### Q1. Mean-Median Score

Calculate the mean-median score associated with this particular district breakdown of votes.

```{r}
votes |> 
    mutate(v_i = d_votes/ (d_votes + r_votes))|>
    summarize(mean_v = mean(v_i), 
    med_v = median(v_i))|>
    mutate(mean_v - med_v)

votes |> 
    mutate(v_i = r_votes/ (d_votes + r_votes))|>
    summarize(mean_v = mean(v_i), 
    med_v = median(v_i))|>
    mutate(mean_v - med_v)
```

### Q2. Efficiency Gap

Calculate the Efficiency Gap (EG) associated with this particular district breakdown of votes.

*Tip*: it may be helpful to write a function that takes a vector of votes across districts for party A and vector of votes for the party B and returns a vector of the number votes wasted by party A across districts.


```{r}

n_wasted <- function(by, against){
    w_threshold <- floor((by +against)/2) +1 
    n_wasted <- ifelse(by > against, by - w_threshold, by)
    n_wasted 
}

votes |>
    mutate(
    d_wasted = n_wasted(d_votes, r_votes), 
    r_wasted = n_wasted(r_votes, d_votes)
    ) |>
    summarize(
        total_d_wasted = sum(d_wasted), 
        total_r_wasted = sum(r_wasted), 
        total_votes = sum(d_votes + r_votes)
    )|>
    mutate(EG = (total_d_wasted - total_r_wasted)/total_votes )


```